### 2012-03-20 docs/NewBCRobustness_paper/paper.tex
committing 3/19 changes
git-svn-id: https://svn.ices.utexas.edu/repos/camellia@381 36274a68-8941-4c93-80ef-affaed5139da

@@

Under the assumption of injectivity of the adjoint of our variational operator $B$, we can define the optimal test norm 
\[
\|v\|_{\rm opt} = \sup_{u\in U}\frac{b(u,v)}{\|u\|_U}
\]
If the solution is in $U$, then choosing $\|v\|=\|v\|_{\rm opt}$ implies 
\[
\|u\|_E = \sup_{v\in V}\frac{b(u,v)}{\|v\|_V} = \sup_{v\in V}\sup_{w\in V}\frac{b(u,v)}{b(w,v)}\|w\|_U =\sup_{w\in V}\frac{b(u,v^*)}{b(w,v^*)}\|w\|_U = \|u\|_U
\]
FIX ABOVE

@@

We consider the model convection-diffusion problem on domain $\Omega$ with boundary $\Gamma$
\begin{align*}
\div (\beta u - \sigma) &= f \\
\frac{1}{\epsilon}\sigma - \grad u &= 0
\end{align*}
with the corresponding variational form 
\[
b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \left(u,\div \tau - \beta \cdot \grad v\right) + \left(\sigma, \epsilon^{-1} \tau + \grad v\right) - \langle \left[\tau_n\right], \widehat{u} \rangle + \langle \widehat{f}_n, \left[v\right] \rangle
\]
where the above inner products are taken element-wise over the finite element mesh $\Omega_h$, and the duality pairings are taken on $\Gamma_h$, the mesh ``skeleton", or union of edges of elements $K$ in $\Omega_h$.  Define the interior skeleton $\Gamma_h^0 = \Gamma_h \setminus \Gamma$.  The functional setting is now well understood as well (see \cite{analysisDPG} for details) - $u,\sigma \in L^2(\Omega)$, $v \in H^1(\Omega_h)$, $\tau \in H({\rm div},\Omega_h)$, where $H^1(\Omega_h)$ and $H({\rm div},\Omega_h)$ are element-wise ``broken" Sobolev spaces.  By duality, $\widehat{u}$ lives in the trace space of $H^1(\Omega_h)$, while $\widehat{f}_n$ comes from the normal trace space of $H({\rm div},\Omega_h)$.

To analyze the robustness of DPG, we note that 
\[
b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \| u \|^2_{L^2}
\]
for the specific continuous test functions $\tau$ and $v$ satisfying
\begin{align*}
\div \tau - \beta \cdot \grad v &= g \\
\frac{1}{\epsilon}\tau + \grad v &= f
\end{align*}
with $g=u$ and $f=0$, and with boundary conditions such that the boundary terms in the bilinear form vanish.  Then, for the above choice of $\left(\tau, v\right)$, 
\begin{align*}
\|u\|_{L^2}^2 &= b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \frac{b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right)}{\|\left(\tau, v\right)\|_V}\|\left(\tau, v\right)\|_V \leq \|u\|_E \|\left(\tau, v\right)\|_V
\end{align*}
We will use the notation $\lesssim$ to denote an $\epsilon$-independent bound.  If we can show $\|\left(\tau, v\right)\|_V \lesssim \| u \|_{L^2}$ for any choice of $u \in L^2$, then dividing through by $\|u\|_{L^2}$ gives
\[
\|u\|_{L^2} \lesssim \|u\|_E
\]
which results in in robust control of the $L^2$ error by the error in the energy norm, in which DPG is optimal.  The study of robustness for the DPG method has now been reduced to proving stability estimates for the adjoint problem.

@@

\subsection{Boundary conditions}

@@

Under the new inflow conditions, the induced bilinear form is now
\[
b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \left(u,\div \tau - \beta \cdot \grad v\right) + \left(\sigma, \epsilon^{-1} \tau + \grad v\right) - \langle \left[\tau_n\right], \widehat{u} \rangle_{\Gamma_-} + \langle \widehat{f}_n, \left[v\right] \rangle_{\Gamma_+}
\]
The corresponding adjoint problem now has boundary conditions
\begin{align}
\tau_n &= 0, \quad x\in \Gamma_-\cup \Gamma_0 \label{eq:adj_bc1}\\
v &= 0, \quad x\in \Gamma_+ \label{eq:adj_bc2}
\end{align}
and it can be shown that DPG is robust using an unweighted version of the test norm in \cite{DPGrobustness}
\[
\|\left(\tau, v\right)\|_{V}^2 = \|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\| + \| \div \tau\| + \frac{1}{\epsilon}\|\tau\|
\]

@@

Additionally, the $\div$ operator is understood now in the weak sense, as the dual operator of $-\grad : H_0^1(\Omega) \rightarrow L^2(\Omega)$, such that $\div f \in \left(H_0^1(\Omega)\right)'$.  $f\cdot n$ is understood as a limit; smooth functions are dense in $H^1(\Omega)$, and we define $f\cdot n$ as the limit of $f_i\cdot n$, where $f_i \in C^\infty(\Omega)$, $f_i \rightarrow f$.

For this analysis, it will be necessary to assume conditions on $\beta$.  For each proof, we require $\beta \in C^2(\bar{\Omega})$ and $\beta, \div \beta = O(1)$.  Additionally, we will assume one or both of the following assumptions
\begin{align}
&\curl \beta = 0, \quad 0<C \leq \left | \beta\right |^2 + \frac{1}{2}\div \beta, \quad C = O(1) \label{a_req}\\
&\grad \beta + \grad \beta ^T - \div \beta I = O(1) \label{b_req}
\end{align}
The second lemma concerns the robust control of $u$ by the energy error; the first and third concern the robust control of field variable $\sigma$ and flux/trace variables, respectively.

@@

\begin{lemma}
\label{lemma_3}
Let $\beta$ satisfy \eqref{a_req} and $\div \beta = 0$, and $f=g=0$.  Then, there holds
\[
\|\grad v\| = \frac{1}{\epsilon}\|\tau\| \lesssim \frac{1}{\epsilon} \| [\tau\cdot n]\|_{\Gamma_h \setminus \Gamma_+} + \frac{1}{\sqrt{\epsilon}} \| [v]\|_{\Gamma_h^0 \cup \Gamma_+}
\]
\end{lemma}
\begin{proof}
We follow \cite{analysisDPG} and \cite{DPGrobustness} in decomposing $\tau$ into $z\in H({\rm curl},\Omega)$ and $\psi\in H^1(\Omega)$ such that
\[
\tau = \left(\epsilon \grad \psi - \beta \psi\right) + \curl z
\]
Specifically, $\psi$ is chosen as the solution to 
\begin{align*}
-\epsilon \grad \psi + \div \left(\beta \psi\right) &= -\div \tau \\
\epsilon \grad \psi \cdot n - \beta_n \psi - \tau\cdot n &= 0, \quad x\in \Gamma_-\\
\psi &= 0, \quad x\in \Gamma_+ 
\end{align*}
Since $\div \beta = 0$, we satisfy \eqref{a_req}, and can use Lemma 1 (with $f = \frac{1}{\epsilon}\tau$) to bound
\[
\epsilon \|\grad \psi\|^2 + \|\psi \|^2 = \frac{1}{\epsilon}\|\tau\|^2
\]
By the above bound and triangle inequality, 
\[
\|\curl z \| \leq \epsilon \|\grad \psi\| + \|\beta \psi\| + \|\tau\| \lesssim \frac{1}{\sqrt{\epsilon}}\|\tau\|
\]
On the other hand, using the decomposition and boundary conditions directly, we have
\begin{align*}
\|\tau\|^2 &= (\tau, \epsilon \grad \psi - \beta \psi + \curl z) = (\tau, \epsilon \grad \psi) - (\tau,\beta \psi) + (\tau,\curl z)  \\
&=(\tau, \epsilon \grad \psi) + \epsilon(\grad v,\beta \psi) - \epsilon(\grad v,\curl z)  \\
&=\epsilon \langle [\tau \cdot n],\psi \rangle - \epsilon  \langle n\cdot \curl z, [v]\rangle = \epsilon \langle [\tau \cdot n],\psi \rangle_{\Gamma_h \setminus \Gamma_+} - \epsilon  \langle n\cdot \curl z, [v]\rangle_{\Gamma_h \cup \Gamma_+}\\
&\lesssim \epsilon \|[\tau \cdot n]\|_{\Gamma_h \setminus \Gamma_+} \|\psi\| + \epsilon \| [v]\|_{\Gamma_h \cup \Gamma_+}\| n\cdot \curl z \| 
\end{align*}
by the boundary norms, which are defined by duality
\begin{align*}
\|[\tau\cdot n]\|_{\Gamma_h} &\coloneqq \sup_{w\in H^1} \frac{\langle [\tau \cdot n], w\rangle}{\|w\|_{H^1(\Omega)}}\\
\|[v]\|_{\Gamma_h} &\coloneqq \sup_{\eta\in H({\rm div})} \frac{\langle v, \eta\cdot n\rangle}{\|\eta\|_{H({\rm div},\Omega)}}
\end{align*}
Applying the above estimates for $\|\psi\|$ and $\|n\cdot \curl z\|$ and noting that $\|\grad v\| = \frac{1}{\epsilon}\|\tau\|$ completes the proof.
\end{proof}
\subsection{Energy norm equivalence (bound from above)}

TODO: FINISH THIS.

@@

In our numerical experiments, we will be interested in the use of the unweighted test norm 
\[
\|\left(\tau, v\right)\|_{V}^2 = \|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\| + \| \div \tau\| + \frac{1}{\epsilon}\|\tau\|
\]
However, the presence of both $\|v\|$ and $\epsilon\|\grad v\|$ terms, and similarly $\|\div \tau\|$ and $\frac{1}{\epsilon}\|\tau\|$ terms, induces boundary layers in the optimal test functions, though only in the crosswind direction.  To avoid this effect, we follow \cite{DPGrobustness} in scaling the $L^2$ contributions of $v$ and $\tau$ such that, when transformed to the reference element, both $v$ and $\grad v$ terms are $O(1)$.  In this paper, we consider only isotropic refinements on quadrilateral elements, such that $h_1 = h_2 = h$, and $|K| = h^2$.
\[
\|\left(\tau, v\right)\|_{V}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\| + \| \div \tau\| + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|
\]
This modified test norm avoids boundary layers, but for adaptive meshes, provides additional stability in areas of heavy refinement.  Intuitively, this is justifiable by the fact that best approximation error is typically largest near boundary layers, where refinements tend to aggregate.  Additionally, test norms that differ from element to element fit cleanly into the DPG framework - as the best approximation error tends to be negligible in areas where the solution is smooth, the finite element error will be small in those areas, even with slightly suboptimal robustness.

@@

For the choice of $\Omega = [0,1]^2$ and $\beta = (1,0)^T$, the convection diffusion equation reduces to
\[
\pd{u}{x} - \epsilon \left(\pdd{u}{x}+ \pdd{u}{y}\right) = f
\]
which, for $f=0$, has an exact solution by separation of variables, allowing us to analyze convergence of DPG for a wide range of $\epsilon$.  For boundary conditions, we impose $u=0$ on $\Gamma_+$ and $\beta_n u - \sigma_n$ on $\Gamma_-$, which reduces to
\begin{align*}
u-\sigma_x &= u_0-\sigma_{x,0}, \quad x=0\\
\sigma_y &=  0, \quad y=0,1\\
\end{align*}
In this case, our exact solution is the series
\[
u(x,y) = C_0 + \sum_{n=1}^\infty C_n \frac{\exp(r_2(x-1)-\exp(r_1(x-1)))}{r_1\exp(-r2) - r_2\exp(-r1)}\cos(n\pi y)
\]
where
\begin{align*}
r_{1,2} &= \frac{1 \pm \sqrt{1 + 4 \epsilon\lambda_n}}{2 \epsilon}\\
\lambda_n &= n^2\pi^2 \epsilon
\end{align*}
The constants $C_n$ depend on a given inflow condition $u_0$ at $x=0$
\[
C_n = \int_0^1 u_0(y) \cos(n\pi y)
\]

@@

\begin{figure}[h]
\centering
\subfigure{
\includegraphics[scale=.37]{figs/wallBC_exact_u.png}
}
\subfigure{
\includegraphics[scale=.37]{figs/wallBC_exact_sigx.png}
}
\subfigure{
\includegraphics[scale=.37]{figs/wallBC_exact_sigy.png}
}
\caption{Solution for $u$, $\sigma_x$, and $\sigma_y$ for $\epsilon = .01$, $C_1 = 1$, $C_n=0$, $n\neq 1$}
\end{figure}

In each case, we begin with a square 4 by 4 mesh of quadrilateral elements with order $p=3$.  We choose $\Delta p = 5$, though we note that the behavior of DPG is nearly identical for any $\Delta p > 3$.  $h$-refinements are executed using a greedy refinement algorithm, where element energy error $e_K^2$ is computed for all elements $K$, and elements such that $e_K^2 \leq \alpha \max_K e_K^2$ are refined.  We make the arbitrary choice of taking $\alpha = .2$ for each of these experiments.

\begin{figure}[h]
\centering
\includegraphics[scale=.4]{figs/u_pointdiff_wallBC.png}
\caption{Adapted mesh and pointwise error for $\epsilon=.01$}
\end{figure}

We are especially interested in the ratio of energy error and $L^2$ error.  Stability estimates imply that, using the above test norm, $\|u-u_h\|_{L^2} / \|u-u_h\|_E \sim C$ independent of $\epsilon$.  Figure~\ref{ratios_simple} seems to imply that, at least for this model problem, $C=O(1)$.

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.38]{figs/errorrates_wallBC.png}
}
\subfigure{
\includegraphics[scale=.38]{figs/L2energyratio_wallBC.png}
}
\caption{$L^2$ and energy errors, and their ratio for $\epsilon=.01$, $\epsilon=.001$, $\epsilon=.0001$}
\label{ratios_simple}
\end{figure}

The effect of the test norm on mesh dependence can be seen in the ratios of $L^2$ to energy error; as the mesh is refined, the constants in front of the $L^2$ terms for $v$ and $\tau$ converge to stationary values, and the ratio of $L^2$ to energy error transitions from a smaller to a larger value.  The transition point happens later for smaller $\epsilon$, which we expect, since the transition of the ratio corresponds to the introduction of elements of order $\epsilon$ through mesh refinement.

@@

In practice, we will not have prior knowledge of $\sigma_n$ at the inflow, and will have to set $\beta_n u - \sigma_n = u_0$, ignoring the viscous contribution to the boundary condition.  The hope is that for small $\epsilon$, this omission will be negligible. Figure~\ref{ratios_noSigma} indicates that, between $\epsilon = .005$ and $\epsilon = .001$, the omission of $\sigma_n$ in the boundary condition becomes negligible, and both our error rates and ratios of $L^2$ to energy error become identical to the case where $\sigma_n$ is explicitly accounted for, well within the range of $\epsilon$ of physical interest. However, even for large $\epsilon = .01$, the $L^2$ error stagnates around $1e-3$, which is only $7%$ relative error.

>>>>

\tableofcontents

@@

Under the assumption of injectivity of the adjoint of our variational operator $B$, we can define the optimal test norm 
\[
\|v\|_{\rm opt} = \sup_{u\in U}\frac{b(u,v)}{\|u\|_U}
\]
If the solution is in $U$, then choosing $\|v\|=\|v\|_{\rm opt}$ implies $\|u\|_E = \|u\|_U$. The proof is simple; we have by definition that $b(u,v) \leq \|u\|_U\|v\|_V$. The reverse $b(u,v) \leq \|u\|_U\|v\|_V$ is proved by
\[
\inf_{u\in U} \frac{\|u\|_E}{\|u\|_U} = \inf_{u\in U} \sup_{v\in V}\frac{b(u,v)}{\|v\|_V\|u\|_U} = \inf_{v\in V} \sup_{u\in U}\frac{b(u,v)}{\|v\|_V\|u\|_U} = \inf_{v\in V} \frac{\|v\|_V}{\|v\|_V} = 1
\]
where we're able to switch the order of $\inf$ and $\sup$ due to the injectivity of $B$ (see  \cite{BabuskaEquivBrezzi}).

@@

We consider the model convection-diffusion problem on domain $\Omega$ with boundary $\Gamma$
\begin{align*}
\div (\beta u - \sigma) &= f \\
\frac{1}{\epsilon}\sigma - \grad u &= 0
\end{align*}
with the corresponding variational form 
\[
b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \left(u,\div \tau - \beta \cdot \grad v\right) + \left(\sigma, \epsilon^{-1} \tau + \grad v\right) - \langle \left[\tau_n\right], \widehat{u} \rangle + \langle \widehat{f}_n, \left[v\right] \rangle
\]
where the above inner products are taken element-wise over the finite element mesh $\Omega_h$, and the duality pairings are taken on $\Gamma_h$, the mesh ``skeleton", or union of edges of elements $K$ in $\Omega_h$.  Define the interior skeleton $\Gamma_h^0 = \Gamma_h \setminus \Gamma$.  The functional setting is now well understood as well (see \cite{analysisDPG} for details) - $u,\sigma \in L^2(\Omega)$, $v \in H^1(\Omega_h)$, $\tau \in H({\rm div},\Omega_h)$, where $H^1(\Omega_h)$ and $H({\rm div},\Omega_h)$ are element-wise ``broken" Sobolev spaces.  By duality, $\widehat{u}$ lives in $H^{1/2}(\Gamma_h)$, the trace space of $H^1(\Omega_h)$, while $\widehat{f}_n$ comes from $H^{-1/2}(\Gamma_h)$, the normal trace space of $H({\rm div},\Omega_h)$.

To analyze the robustness of DPG, we note that 
\[
b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \| u \|^2_{L^2}
\]
for the specific continuous test functions $\tau$ and $v$ satisfying
\begin{align*}
\div \tau - \beta \cdot \grad v &= g \\
\frac{1}{\epsilon}\tau + \grad v &= f
\end{align*}
with $g=u$ and $f=0$, and with boundary conditions such that the boundary terms in the bilinear form vanish.  Then, for the above choice of $\left(\tau, v\right)$, 
\begin{align*}
\|u\|_{L^2}^2 &= b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \frac{b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right)}{\|\left(\tau, v\right)\|_V}\|\left(\tau, v\right)\|_V \leq \|u\|_E \|\left(\tau, v\right)\|_V
\end{align*}
We will use the notation $\lesssim$ to denote an $\epsilon$-independent bound.  If we can show $\|\left(\tau, v\right)\|_V \lesssim \| g \|_{L^2}$ for any choice of $g \in L^2$, then, for $g=u$, dividing through by $\|u\|_{L^2}$ gives
\[
\|u\|_{L^2} \lesssim \|u\|_E
\]
which results in in robust control of the $L^2$ error by the error in the energy norm, in which DPG is optimal.  The study of robustness for the DPG method has now been reduced to proving stability estimates for the adjoint problem.

@@

\subsection{Boundary conditions and adjoint stability}

@@

Under the new inflow conditions, the induced bilinear form is now
\[
b\left(\left(u,\sigma, \widehat{u}, \widehat{\sigma}_n\right), \left(\tau, v\right)\right) = \left(u,\div \tau - \beta \cdot \grad v\right) + \left(\sigma, \epsilon^{-1} \tau + \grad v\right) - \langle \left[\tau_n\right], \widehat{u} \rangle_{\Gamma_-} + \langle \widehat{f}_n, \left[v\right] \rangle_{\Gamma_+}
\]
The corresponding adjoint problem for $L^2$ stability of $u$ now has boundary conditions
\begin{align}
\tau_n &= 0, \quad x\in \Gamma_-\cup \Gamma_0 \label{eq:adj_bc1}\\
v &= 0, \quad x\in \Gamma_+ \label{eq:adj_bc2}
\end{align}

\subsection{Norms on $U$}

Given the boundary conditions on $\widehat{u}$ and $\widehat{f}_n$, we can now also define the trace norms as (canonical) minimum energy extension norms 
\begin{align*}
\|\widehat{u}\| &= \inf_{w\in H^1(\Omega), \left.w\right|_{\Gamma_+}=0, \left.w\right |_{\Gamma_h\setminus \Gamma_+ } = \widehat{u}} \|w\|_{H^1(\Omega)}\\
\|\widehat{f}_n\| &= \inf_{q\in H({\rm div},\Omega), \left.q\cdot n\right|_{\Gamma_-}= 0, \left. q\cdot n \right |_{\Gamma_h} = \widehat{f}_n} \|q\|_{H({\rm div},\Omega)}
\end{align*}
We will construct norms on $U$ as combinations of norms on each variable
\[
\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)\|_U = C_1 \|u\|_{L^2} + C_2 \|\sigma\|_{L^2} + C_3 \|\widehat{u}\| + C_4 \|\widehat{f}_n\|
\]

\subsection{Main result}

For this analysis, it will be necessary to assume certain conditions on $\beta$.  For each proof, we require $\beta \in C^2(\bar{\Omega})$ and $\beta, \div \beta = O(1)$.  Additionally, we will assume one or both of the following assumptions
\begin{align}
&\curl \beta = 0, \quad 0<C \leq \left | \beta\right |^2 + \frac{1}{2}\div \beta, \quad C = O(1) \label{a_req}\\
&\grad \beta + \grad \beta ^T - \div \beta I = O(1) \label{b_req}\\
&\div \beta = 0 \label{c_req}
\end{align}

The primary result of this paper is that, if conditions \ref{a_req},\ref{b_req}, and \ref{c_req} are all satisfied, the DPG method provides robust control over the $L^2$ error 
\[
\left\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)-\left(u_h,\sigma_h,\widehat{u}_h,\widehat{f}_{n,h}\right)\right\|_{U_1} \lesssim \inf_{w\in U_h}\left\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)-w_h \right\|_{E}
\]
using the unweighted version of the test norm in \cite{DPGrobustness}
\[
\|\left(\tau, v\right)\|_{V}^2 = \|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\| + \| \div \tau\| + \frac{1}{\epsilon}\|\tau\|
\]
where $\|\cdot \|_{U_1}$ is defined as
\[
\left\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)\right\|_{U_1}^2 \coloneqq \|u\|^2 + \|\sigma\|^2 + \epsilon^2\|\widehat{u}\|^2+\epsilon \|\widehat{f}_n\|^2.
\]
A second, weaker result is an upper bound on the energy norm
\[
\left\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)-\left(u_h,\sigma_h,\widehat{u}_h,\widehat{f}_{n,h}\right)\right\|_{E} \lesssim \inf_{w\in U_h}\left\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)-w_h\right\|_{U_2}
\]
where $\|\cdot \|_{U_2}$ is defined as
\[
\left\|\left(u,\sigma,\widehat{u},\widehat{f}_n\right)\right\|_{U_2}^2 \coloneqq \|u\|^2 + \frac{1}{\epsilon^2}\|\sigma\|^2 + \frac{1}{\epsilon}\|\widehat{u}\|^2+\frac{1}{\epsilon} \|\widehat{f}_n\|^2.
\]
The upper bound on the energy norm appears to depend on $\epsilon^{-1}$, implying that, at most, we have the bound from below $\|u-u_h\|_{U} \geq \epsilon\|u-u_h\|_E$. However, numerical experiments demonstrate much better results, and suggest that, at least for our examples, $\|u-u_h\|_E \lesssim \|u-u_h\|_{U}$, implying that the induced energy norm is equivalent to the canonical norms we've chosen on our group variable $\left(u,\sigma,\widehat{u},\widehat{f}_n\right)$.

@@

Additionally, the $\div$ operator is understood now in the weak sense, as the dual operator of $-\grad : H_0^1(\Omega) \rightarrow L^2(\Omega)$, such that $\div f \in \left(H_0^1(\Omega)\right)'$. The normal trace $f\cdot n$ is understood as a limit; smooth functions are dense in $H^1(\Omega)$, and we define $f\cdot n$ as the limit of $f_i\cdot n$, where $f_i \in C^\infty(\Omega)$, $f_i \rightarrow f$.

The second lemma concerns the robust control of $u$ by the energy error; the first and third concern the robust control of field variable $\sigma$ and flux/trace variables, respectively.

@@

\begin{lemma}
\label{lemma_3}
Let $\beta$ satisfy \eqref{a_req} and \eqref{c_req}, and let $f=g=0$.  Then, there holds
\[
\|\grad v\| = \frac{1}{\epsilon}\|\tau\| \lesssim \frac{1}{\epsilon} \| [\tau\cdot n]\|_{\Gamma_h \setminus \Gamma_+} + \frac{1}{\sqrt{\epsilon}} \| [v]\|_{\Gamma_h^0 \cup \Gamma_+}
\]
\end{lemma}
\begin{proof}
We follow \cite{analysisDPG} and \cite{DPGrobustness} in decomposing $\tau$ into $z\in H({\rm curl},\Omega)$ and $\psi\in H^1(\Omega)$ such that
\[
\tau = \left(\epsilon \grad \psi - \beta \psi\right) + \curl z
\]
Specifically, $\psi$ is chosen as the solution to 
\begin{align*}
-\epsilon \grad \psi + \div \left(\beta \psi\right) &= -\div \tau \\
\epsilon \grad \psi \cdot n - \beta_n \psi - \tau\cdot n &= 0, \quad x\in \Gamma_-\\
\psi &= 0, \quad x\in \Gamma_+ 
\end{align*}
Since $\div \beta = 0$, we satisfy \eqref{a_req}, and can use Lemma 1 (with $f = \frac{1}{\epsilon}\tau$) to bound
\[
\epsilon \|\grad \psi\|^2 + \|\psi \|^2 = \frac{1}{\epsilon}\|\tau\|^2
\]
By the above bound and triangle inequality, 
\[
\|\curl z \| \leq \epsilon \|\grad \psi\| + \|\beta \psi\| + \|\tau\| \lesssim \frac{1}{\sqrt{\epsilon}}\|\tau\|
\]
On the other hand, using the decomposition and boundary conditions directly, we have
\begin{align*}
\|\tau\|^2 &= (\tau, \epsilon \grad \psi - \beta \psi + \curl z) = (\tau, \epsilon \grad \psi) - (\tau,\beta \psi) + (\tau,\curl z)  \\
&=(\tau, \epsilon \grad \psi) + \epsilon(\grad v,\beta \psi) - \epsilon(\grad v,\curl z)  \\
&=\epsilon \langle [\tau \cdot n],\psi \rangle - \epsilon  \langle n\cdot \curl z, [v]\rangle = \epsilon \langle [\tau \cdot n],\psi \rangle_{\Gamma_h \setminus \Gamma_+} - \epsilon  \langle n\cdot \curl z, [v]\rangle_{\Gamma_h \cup \Gamma_+}\\
&\lesssim \epsilon \|[\tau \cdot n]\|_{\Gamma_h \setminus \Gamma_+} \|\psi\| + \epsilon \| [v]\|_{\Gamma_h \cup \Gamma_+}\| n\cdot \curl z \| 
\end{align*}
by the boundary norms for $v$ and $\tau$, which are defined by duality
\begin{align*}
\|[\tau\cdot n]\|_{\Gamma_h} &\coloneqq \sup_{w\in H^1} \frac{\langle [\tau \cdot n], w\rangle}{\|w\|_{H^1(\Omega)}}\\
\|[v]\|_{\Gamma_h} &\coloneqq \sup_{\eta\in H({\rm div})} \frac{\langle v, \eta\cdot n\rangle}{\|\eta\|_{H({\rm div},\Omega)}}
\end{align*}
Applying the above estimates for $\|\psi\|$ and $\|n\cdot \curl z\|$ and noting that $\|\grad v\| = \frac{1}{\epsilon}\|\tau\|$ completes the proof.
\end{proof}

\subsection{Energy norm equivalence (bound from above)}

We've shown the robust bound of a norm on $U$ by energy error; for a full equivalence statement between the energy norm and norms on $U$, we require a bound on the energy norm by some norm on $U$.

@@

In our numerical experiments, we will be interested in the use of the unweighted test norm 
\[
\|\left(\tau, v\right)\|_{V}^2 = \|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\| + \| \div \tau\| + \frac{1}{\epsilon}\|\tau\|
\]
However, the presence of both $\|v\|$ and $\epsilon\|\grad v\|$ terms, and similarly $\|\div \tau\|$ and $\frac{1}{\epsilon}\|\tau\|$ terms, induces boundary layers in the optimal test functions, though only in the crosswind direction.  To avoid this effect, we follow \cite{DPGrobustness} in scaling the $L^2$ contributions of $v$ and $\tau$ such that, when transformed to the reference element, both $v$ and $\grad v$ terms are $O(1)$.  In this paper, we consider only isotropic refinements on quadrilateral elements, such that $h_1 = h_2 = h$, and $|K| = h^2$. Our test norm for an element $K$ is now
\[
\|\left(\tau, v\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\| + \epsilon \|\grad v\| + \|\beta \cdot \grad v\| + \| \div \tau\| + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|
\]
This modified test norm avoids boundary layers, but for adaptive meshes, provides additional stability in areas of heavy refinement.  Intuitively, this is justifiable by the fact that best approximation error is typically largest near boundary layers, where refinements tend to aggregate.  %Additionally, test norms that differ from element to element fit cleanly into the DPG framework - as the best approximation error tends to be negligible in areas where the solution is , the finite element error will be small in those areas, even with slightly suboptimal robustness.

@@

For the choice of $\Omega = (0,1)^2$ and $\beta = (1,0)^T$, the convection diffusion equation reduces to
\[
\pd{u}{x} - \epsilon \left(\pdd{u}{x}+ \pdd{u}{y}\right) = f
\]
which, for $f=0$, has an exact solution by separation of variables, allowing us to analyze convergence of DPG for a wide range of $\epsilon$.  For boundary conditions, we impose $u=0$ on $\Gamma_+$ and $\beta_n u - \sigma_n$ on $\Gamma_-$, which reduces to
\begin{align*}
u-\sigma_x &= u_0-\sigma_{x,0}, \quad x=0\\
\sigma_y &=  0, \quad y=0,1\\
\end{align*}
In this case, our exact solution is the series
\[
u(x,y) = C_0 + \sum_{n=1}^\infty C_n \frac{\exp(r_2(x-1)-\exp(r_1(x-1)))}{r_1\exp(-r2) - r_2\exp(-r1)}\cos(n\pi y)
\]
where
\begin{align*}
r_{1,2} &= \frac{1 \pm \sqrt{1 + 4 \epsilon\lambda_n}}{2 \epsilon}\\
\lambda_n &= n^2\pi^2 \epsilon
\end{align*}
The constants $C_n$ depend on a given inflow condition $u_0$ at $x=0$
\[
C_n = \int_0^1 u_0(y) \cos(n\pi y)
\]

@@

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.37]{figs/wallBC_exact_u.png}
}
\subfigure{
\includegraphics[scale=.37]{figs/wallBC_exact_sigx.png}
}
\subfigure{
\includegraphics[scale=.37]{figs/wallBC_exact_sigy.png}
}
\caption{Solution for $u$, $\sigma_x$, and $\sigma_y$ for $\epsilon = .01$, $C_1 = 1$, $C_n=0$, $n\neq 1$}
\end{figure}
In each case, we begin with a square 4 by 4 mesh of quadrilateral elements with order $p=3$.  We choose $\Delta p = 5$, though we note that the behavior of DPG is nearly identical for any $\Delta p > 3$.  $h$-refinements are executed using a greedy refinement algorithm, where element energy error $e_K^2$ is computed for all elements $K$, and elements such that $e_K^2 \leq \alpha \max_K e_K^2$ are refined.  We make the arbitrary choice of taking $\alpha = .2$ for each of these experiments.

\begin{figure}[h!]
\centering
\includegraphics[scale=.4]{figs/u_pointdiff_wallBC.png}
\caption{Adapted mesh and pointwise error for $\epsilon=.01$}
\end{figure}
We are especially interested in the ratio of energy error and total $L^2$ error in both $\sigma$ and $u$.  Stability estimates imply that, using the above test norm, $\|u-u_h\|_{L^2} / \|u-u_h\|_E \sim C$ independent of $\epsilon$.  Figure~\ref{ratios_simple} seems to imply that, at least for this model problem, $C=O(1)$.

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.38]{figs/errorrates_wallBC.png}
}
\subfigure{
\includegraphics[scale=.38]{figs/L2energyratio_wallBC.png}
}
\caption{$L^2$ and energy errors, and their ratio for $\epsilon=.01$, $\epsilon=.001$, $\epsilon=.0001$}
\label{ratios_simple}
\end{figure}
The effect of a mesh dependent test norm can be seen in the ratios of $L^2$ to energy error; as the mesh is refined, the constants in front of the $L^2$ terms for $v$ and $\tau$ converge to stationary values, and the ratio of $L^2$ to energy error transitions from a smaller to a larger value.  The transition point happens later for smaller $\epsilon$, which we expect, since the transition of the ratio corresponds to the introduction of elements of order $\epsilon$ through mesh refinement.

@@

In practice, we will not have prior knowledge of $\sigma_n$ at the inflow, and will have to set $\beta_n u - \sigma_n = u_0$, ignoring the viscous contribution to the boundary condition.  The hope is that for small $\epsilon$, this omission will be negligible. Figure~\ref{ratios_noSigma} indicates that, between $\epsilon = .005$ and $\epsilon = .001$, the omission of $\sigma_n$ in the boundary condition becomes negligible, and both our error rates and ratios of $L^2$ to energy error become identical to the case where $\sigma_n$ is explicitly accounted for, well within the range of $\epsilon$ of physical interest. For large $\epsilon = .01$, the $L^2$ error stagnates around $1e-3$, or about $7\%$ relative error.

\begin{figure}[h!]
\centering
\subfigure{
\includegraphics[scale=.38]{figs/rates_noSigma.png}
}
\subfigure{
\includegraphics[scale=.38]{figs/ratio_noSigma.png}
}
\caption{$L^2$ and energy errors and their ratio when neglecting $\sigma_n$ at the inflow.}
\label{ratios_noSigma}
\end{figure}
